---
title: "Cluster-achilles"
author: "Carla Cohen"
date: "`r Sys.Date()`"
output: html_document
---

Read in filtered RDS objects
Perform PCA & visualise
Perform clustering & visualise


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(Seurat)
library(biomaRt)
library(scater)

```

Read in RDS files after filtering


```{r, echo=FALSE}
#make a list of sample files
file_list <- list.files("RDS_objects_test.dir/", pattern="_filtered_SeuratObject.rds")

#Make a new list for the seurat objects
so <- list()

# generate a list of Seurat objects
for (i in 1:length(file_list)){
    
    so[[i]] <- readRDS(paste0("RDS_objects_test.dir/", file_list[i]))
    
}

```


Perform SCTransform

```{r}

for (i in 1:length(file_list)){
    
    so[[i]] <- SCTransform(so[[i]],
                           vars.to.regress = "percent_mt")
    }

```
Have  alook a the SCTransform output

```{r, eval = FALSE, echo=FALSE}

for (i in 1:length(file_list)){
    print(so[[i]])
    print(so[[i]][[]] %>% head()) #should see new columns in metadata
    print(dim(so[[i]]$RNA))
    print(dim(so[[i]]$SCT))
    
    }



```

```{r}
so[[1]]@assays$RNA@meta.features %>% head()

so_unfiltered_test <- readRDS("RDS_objects.dir/OMB1250-Ach-MB1_unfiltered_SeuratObject.rds")
so_unfiltered_test@assays$RNA@meta.features %>% head()


```

NB this section can be removed once fixed in QC-1 and re-run
Access biomaRT to get a mapping of EnsemblIDs to gene names

```{r, echo=FALSE}

# Access the useMart database
# useMart keeps not working every so often (issue https://github.com/grimbough/biomaRt/issues/31)
ensembl <- try(useMart("ensembl", dataset = "hsapiens_gene_ensembl"))
if(class(ensembl) == "try-error"){
  httr::set_config(httr::config(ssl_verifypeer = FALSE))
  ensembl <- useMart("ensembl", dataset = "hsapiens_gene_ensembl")
}

# get a df of ensemble_id vs gene symbol
mapping <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"), mart = ensembl)
```

### Add gene symbols to the meta.features of the RNA assay

Where no gene symbol is available, leave in the EnsemblID  
For the Seurat object this has to be added for each assay.  
For Bioconductor SingleCellExperiment this can be added as rowData.  


```{r, echo=FALSE, message = FALSE}

df <- list()
for(i in 1:length(so)){
    
    #make a df with ensemble_id and hgnc_symbol
    df[[i]] <- data.frame("ensembl_gene_id" = rownames(so[[i]])) %>% 
        left_join(mapping)
    
    
    #add a column of unique names, remove the "_" from ensembl_ids
    df[[i]]$unique_name <- uniquifyFeatureNames(df[[i]]$ensembl_gene_id, df[[i]]$hgnc_symbol)
    df[[i]]$unique_name <- gsub("^_", "", df[[i]]$unique_name)
    df[[i]] <- df[[i]] %>% distinct(ensembl_gene_id, .keep_all = TRUE) #remove duplicated rows
    print(head(df[[i]]))
    
    #add ensembl_IDs to row names
    rownames(df[[i]]) <- rownames(so[[i]])
    
    #add to the meta.features in the RNA assay
    so[[i]][["RNA"]]@meta.features <- df[[i]]
    print(head(so[[i]][["RNA"]]@meta.features))
    
}

```

? can we transfer the names from RNA to SCT assay?

Plot variable features

```{r}


plot_list <- list()

for (i in 1:length(so)){
    
    top10 <- head(VariableFeatures(so[[i]]), 10)
    
    plot1 <- VariableFeaturePlot(so[[i]])
  
  # label with gene names
  plot_list[[i]] <- LabelPoints(plot = plot1, points = top10, #labels = top10_genes$hgnc_symbol, 
                       repel = TRUE, xnudge = 0, ynudge = 0)

  print(plot_list[[i]])

    
    }


```



Visualise VlnPlots before and afer SCTransform

```{r}

plot_list <- list()

for (i in 1:length(so)) {
    
    # do a violin plot
    plot_list[[i]] <- VlnPlot(so[[i]], c("nCount_RNA", "nCount_SCT"), pt.size = 0)+
        theme(legend.position="none")
    print(plot_list[[i]])
}

```

? Do we need to add gene  names to the SCT assay?


Run PCA and do the Elbow plot to check how many dimensions we should use going forward


```{r}

plot_list <- list()

for (i in 1:length(so)) {
    
    #Calculate PCA
    so[[i]] <- RunPCA(so[[i]], reduction.name = "sct.pca")
    
    #do elbow plot
    plot_list[[i]] <- ElbowPlot(so[[i]], ndims = 50)
    print(plot_list[[i]])
}



```

Run tSNE and UMAP on the selected number of PCs. 
Might this be different for different samples?


```{r}
for (i in 1:length(so)) {
    
    so[[i]] <- RunUMAP(seurat_after_qc,
                       reduction = "sct.pca",
                       dims = 1:20,
                       reduction.name = "sct.umap")
    so[[i]] <- RunTSNE()
    
}
```




From filter-achilles.Rmd

Perform denoise on the filtered object, then recalculate the UMAP on the reduced number of PCs. 

```{r, echo = FALSE}
sce_decontX_doublet_discard_filtered_denoise <- list()
df <- data.frame()

for (i in 1:length(sce_decontX_doublet_discard_filtered)){
    
    sce_decontX_doublet_discard_filtered_denoise[[i]] <-
        scran::denoisePCA(sce_decontX_doublet_discard_filtered[[i]],
                          dec_filtered[[i]], 
                          subset.row=hvg_filtered[[i]])
    
    sample <- project_title[[i]]
    denoise_PCs <- ncol(reducedDim(sce_decontX_doublet_discard_filtered_denoise[[i]], "PCA"))
    
    df[i,1] <- sample
    df[i, 2] <- denoise_PCs
    
    sce_decontX_doublet_discard_filtered_denoise[[i]] <- scater::runUMAP(sce_decontX_doublet_discard_filtered_denoise[[i]], 
                                                                 dimred = 'PCA', 
                                                                 pca = denoise_PCs,
                                                                 external_neighbors=TRUE)
    
}
    
colnames(df) <- c("sample", "number of PCs")

write.table (df, "Files/denoise_PCs.txt", row.names = FALSE, sep = "\t")

df

```

Plot the UMAPs side by side, with and without denoise.

```{r, fig.height = 12, message = FALSE}
plot_list <- list()

for (i in 1:length(sce_decontX_doublet_discard_filtered)){
    
    p1 <- plotReducedDim(sce_decontX_doublet_discard_filtered[[i]], dimred = "UMAP", colour_by = "sex")+
        scale_colour_viridis_d(begin = 0, end = 0.75)+
        ggtitle("Before denoise")+
        theme(legend.position="none")
    
    p2 <- plotReducedDim(sce_decontX_doublet_discard_filtered_denoise[[i]], dimred = "UMAP", colour_by = "sex")+
        scale_colour_viridis_d(begin = 0, end = 0.75)+
        ggtitle("After denoise")+
        theme(legend.position="none")

        p3 <- plot_grid(p1, p2, nrow = 1)
    title <- ggdraw() + draw_label(project_title[[i]], fontface='bold', size = 16)
    plot_list[[i]] <- plot_grid(title, p3, ncol=1, rel_heights=c(0.1, 1))
}


#add an overall title and arrange the graphs on one page
title <- ggdraw() + draw_label("Denoise PCA", fontface='bold', size = 24)
p <- plot_grid(plotlist = plot_list, label_size = 24, ncol = 2)
p <- plot_grid(title, p, ncol=1, rel_heights=c(0.1, 1))

p
    


```

```{r, include=FALSE}
#save the plot
ggsave("Filtered_Figures.dir/Denoise_UMAP.png", p, device = "png", width = 10, height = 14, bg = "white")
ggsave("Filtered_Figures.dir/Denoise_UMAP.pdf", p, device = "pdf", width = 12, height = 24, bg = "white")
```
