---
title: "Cluster-achilles"
author: "Carla Cohen"
date: "`r Sys.Date()`"
output: html_document
---


This workflow is part of the scflow seurat pipeline and performs dimensionality reduction and clustering of individual samples. 

The following parameters can be specified in the yml file:
* Files to remove (e.g. that have failed QC)  
* Normalisation method (SCTransform, SCTransform v2, Log Normalisation)  
* File format (e.g. png, pdf, svg)  
* Clustering resoultion. The script calculates resolutions from 0.1-1.0 in 0.1 increments. The default is 0.5.  


Steps in the workflow  
*  Read in filtered RDS objects  
*  Convert EnsemblIDs to gene names in RNA  assay
*  Perform Normalisation using preferred method, followed by RunPCA, FindNeighbours and RunUMAP. Default parameters are used here, these will be optimised later. 
* Perform clustering at multiple resolutions
*  Convert EnsemblIDs to gene names  
* Save the RDS files in /RDS_objects.dir/clustered/{norm method}
*  Visualise variable genes, Elbow Plot, clustree and UMAPs  
  



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 10, fig.height = 16) #set optimal fig.width and fig.height here

library(tidyverse)
library(Seurat)
library(scater)
library(cowplot)
library(glmGamPoi)
library(clustree)
library(yaml)
library(viridis)

# Read in the yml file
ini <- read_yaml("pipeline.yml")

```

## Import the data

Make a list of files
Remove some samples due to them failing QC
MSK1250-Ach-MB2
MSK1284 all samples


```{r}

#make a list of sample files
file_list <- list.files("RDS_objects.dir/filtered", pattern="discard_filtered_SeuratObject.rds")
names(file_list) <- str_replace(file_list, "_decontX_doublet_discard_filtered_SeuratObject.rds", "")

cat ("File list before sample removal")
cat ("\n")
names(file_list)


# get the positions of unwanted files in the list
positions <- match(ini$files_to_remove, names(file_list))

# remove certain files using those names
file_list <- file_list[-c(positions)] 


cat ("\n")
cat ("File list after sample removal")
cat ("\n")
names(file_list)


```


Read in RDS files after filtering


```{r, echo=FALSE}

#Make a new list for the seurat objects
so <- list()

# generate a list of Seurat objects
for (i in 1:length(file_list)){
    
    so[[i]] <- readRDS(paste0("RDS_objects.dir/filtered/", file_list[i]))
    #set the project name (should be done in filtering workflow)
    so[[i]]@project.name <- str_replace(file_list[[i]], "_decontX_doublet_discard_filtered_SeuratObject.rds", "")
}

```


Convert rownames of RNA assay data & counts slots to unique names


```{r, include=FALSE}

for (i in 1:length(so)){
  
  # add gene names to RNA assay
  df <- left_join(data.frame("ensembl_gene_id" = rownames(so[[i]]@assays$RNA@data)), so[[i]]@assays$RNA@meta.features)
  rownames(so[[i]]@assays$RNA@data) <- df$unique_name
  rownames(so[[i]]@assays$RNA@counts) <- df$unique_name

}

```


### Parameters
Extract some names that depend on the normalisation method performed

```{r}

if (ini$normalisation_method == "SCTransform"){
        
        # save the reduction names
        pca_name <- "sct.pca"
        umap_name <- "sct.umap"
        
        # save the resolution prefix
        res_prefix <- "SCT_snn_res."
        
    } else if (ini$normalisation_method == "SCTransformv2"){
    
    # save the reduction names
     pca_name <- "sctv2.pca"
     umap_name <- "sctv2.umap"
     
     # save the resolution prefix
     res_prefix <- "SCTv2_snn_res."
    
    } else if (ini$normalisation_method == "LogNormalise"){
    
     # save the reduction names
     pca_name <- "lognorm.pca"
     umap_name <- "lognorm.umap"
     
     # save the resolution prefix
     res_prefix <- "RNA_snn_res."
        
}
    
print(ini$normalisation_method) 
print(pca_name)
print(umap_name)
print(res_prefix)

```



## Perform Normalisation 

Default method is SCTransform, can also update to logNormalise or SCTransform v2 using the yml file. 
Perform the downstream steps with default parameters and using 30 PCs (this will be optimised later)


```{r, message=FALSE}

for (i in 1:length(so)){
    
    cat ("Analysing sample", i, ":", so[[i]]@project.name)
    cat ("\n")
    
    if (ini$normalisation_method == "SCTransform"){
        
        so[[i]] <- SCTransform(so[[i]],
                           method = "glmGamPoi", #added to increase efficiency
                           vars.to.regress = "percent_mt") %>% 
        RunPCA(reduction.name = "sct.pca") %>% 
        FindNeighbors(dims = 1:ini$n_dims, reduction = "sct.pca") %>%
        RunUMAP(dims = 1:ini$n_dims, reduction = "sct.pca", reduction.name = "sct.umap") 
    
        # add gene names to SCT assay
        df <- left_join(data.frame("ensembl_gene_id" = rownames(so[[i]]@assays$SCT@data)), so[[i]]@assays$RNA@meta.features)
        rownames(so[[i]]@assays$SCT@data) <- df$unique_name
        rownames(so[[i]]@assays$SCT@counts) <- df$unique_name
        
    }
    
    else if (ini$normalisation_method == "SCTransformv2"){
    
    so[[i]] <- SCTransform(so[[i]],
                           method = "glmGamPoi", #added to increase efficiency
                           vars.to.regress = "percent_mt", 
                           vst.flavor = "v2", 
                           new.assay.name = "SCTv2") %>% # using SCTransform v2
        RunPCA(reduction.name = "sctv2.pca") %>% 
        FindNeighbors(dims = 1:ini$n_dims, reduction = "sctv2.pca") %>%
        RunUMAP(dims = 1:ini$n_dims, reduction = "sctv2.pca", reduction.name = "sctv2.umap") 
    
      # add gene names to SCTv2 assay
        df <- left_join(data.frame("ensembl_gene_id" = rownames(so[[i]]@assays$SCTv2@data)), so[[i]]@assays$RNA@meta.features)
        rownames(so[[i]]@assays$SCTv2@data) <- df$unique_name
        rownames(so[[i]]@assays$SCTv2@counts) <- df$unique_name
        
   }
    
    else if (ini$normalisation_method == "LogNormalise"){
        so[[i]] <- so[[i]] %>% 
           NormalizeData() %>%
           FindVariableFeatures() %>% 
           ScaleData() %>%
           RunPCA(reduction.name = "lognorm.pca")%>%
           FindNeighbors(dims = 1:ini$n_dims, reduction = "lognorm.pca") %>%
           RunUMAP(dims = 1:ini$n_dims, reduction = "lognorm.pca", reduction.name = "lognorm.umap") 
        
    }
    
  
}

 

```


### Add multiple resolutions of clustering
Calculate clusters at resolutions 0.1 to 1 with 0.1 intervals.

```{r, message=FALSE}

so <- lapply(so, FindClusters, resolution = seq(0.1, 1, 0.1))


```
### Save the RDS objects

```{r}
dir.create ("RDS_objects.dir/clustered/", showWarnings = FALSE)
dir.create (paste0("RDS_objects.dir/clustered/", ini$normalisation_method), showWarnings = FALSE)

for(i in 1:length(so)){
    
    saveRDS(so[[i]], paste0("RDS_objects.dir/clustered/", ini$normalisation_method, "/", so[[i]]@project.name, "_filtered_clustered_SeuratObject.rds"))
    
}

```



### Plot the variable features following normalisation


```{r, fig.width=12, fig.height=18, message=FALSE}
plot_list <- list()

for (i in 1:length(so)){
    
    
    p1 <- VariableFeaturePlot(so[[i]], pt.size = 0.2)+
      scale_colour_viridis_d(begin = 0, end = 0.75)
    
    plot_list[[i]] <- LabelPoints(plot = p1, points = VariableFeatures(so[[i]])[1:10], 
                      repel = TRUE, xnudge = 0, ynudge = 0)+
        ggtitle(so[[i]]@project.name)
    
}

title <- ggdraw() + draw_label(paste0(ini$normalisation_method, " variable features"), fontface='bold', size = 24)
p <- plot_grid(plotlist = plot_list, label_size = 12, ncol = 2)
p <- plot_grid(title, p, ncol=1, rel_heights=c(0.05, 1))

p


```

```{r, include=FALSE}
#save the plot
dir.create(paste0("Clustering_Figures.dir/", ini$normalisation_method), showWarnings = FALSE)
ggsave(paste0("Clustering_Figures.dir/", ini$normalisation_method, "/HVG.", ini$file_format), p, device = ini$file_format, width = 10, height = 18, bg = "white")


```



### ElbowPlot

```{r, fig.width=10, fig.height = 18, message=FALSE}
plot_list <- list()

for (i in 1:length(so)){
    
    plot_list[[i]] <- ElbowPlot(so[[i]], reduction = pca_name, ndims = 50)+ggtitle(so[[i]]@project.name, )
  
}

title <- ggdraw() + draw_label(paste0(ini$normalisation_method, " Elbow Plot"), fontface='bold', size = 24)
p <- plot_grid(plotlist = plot_list, label_size = 12, ncol = 2)
p <- plot_grid(title, p, ncol=1, rel_heights=c(0.05, 1))

p


```


```{r, include=FALSE}
#save the plot
ggsave(paste0("Clustering_Figures.dir/", ini$normalisation_method, "/Elbow_plot.", ini$file_format), p, device = ini$file_format, width = 10, height = 18, bg = "white")


```


### Clustree

```{r, fig.width=12, fig.height=20}

plot_list <- list()

for (i in 1:length(so)){
    
    plot_list[[i]] <- clustree(so[[i]], prefix = res_prefix) + 
        theme(legend.position="none")+
        ggtitle(so[[i]]@project.name)
    
}

title <- ggdraw() + draw_label(paste0(ini$normalisation_method, " Clustree Plot"), fontface='bold', size = 24)
p <- plot_grid(plotlist = plot_list, label_size = 12, ncol = 2)
p <- plot_grid(title, p, ncol=1, rel_heights=c(0.1, 1))

p



```


```{r, include=FALSE}
#save the plot
ggsave(paste0("Clustering_Figures.dir/", ini$normalisation_method, "/Clustree_plot.", ini$file_format), p, device = ini$file_format, width = 12, height = 20, bg = "white")


```

### Plot UMAP with clustering

```{r, fig.width=10, fig.height = 18,message=FALSE}
plot_list <- list()

for (i in 1:length(so)){
    
    # set chosen resolution (default = 0.5)
    
    so[[i]][["seurat_clusters"]] <- so[[i]][[paste0(res_prefix, ini$resolution)]]
    
    # plot UMAP
    plot_list[[i]] <- DimPlot(so[[i]], reduction = umap_name, label = TRUE, repel = TRUE) +
        theme(legend.position="none")+
        ggtitle(so[[i]]@project.name)
    
}

title <- ggdraw() + draw_label(paste0(ini$normalisation_method, " Clustering at resolution = ", ini$resolution), fontface='bold', size = 24)
p <- plot_grid(plotlist = plot_list, label_size = 12, ncol = 2)
p <- plot_grid(title, p, ncol=1, rel_heights=c(0.05, 1))

p


```

```{r, include=FALSE}
#save the plot
ggsave(paste0("Clustering_Figures.dir/", ini$normalisation_method, "/UMAP_plot.", ini$file_format), p, device = ini$file_format, width = 12, height = 18, bg = "white")


```

### Feature plot for mitochondrial content

```{r, fig.width=10, fig.height = 18, message = FALSE}

plot_list <- list()

for (i in 1:length(so)){
    
    plot_list[[i]]  <- FeaturePlot(so[[i]], features = "percent_mt", reduction = umap_name)+
        scale_colour_gradientn(colours = viridis(256, option = "D"), name = "percent_mt", limits = c(0, 25))+
        ggtitle(so[[i]]@project.name)

}

title <- ggdraw() + draw_label(paste0(ini$normalisation_method, " Percent MT genes on UMAP"), fontface='bold', size = 24)
p <- plot_grid(plotlist = plot_list, label_size = 12, ncol = 2)
p <- plot_grid(title, p, ncol=1, rel_heights=c(0.1, 1))

p


```


```{r, include=FALSE}
#save the plot
ggsave(paste0("Clustering_Figures.dir/", ini$normalisation_method, "/UMAP_percent_mt.", ini$file_format), p, device = ini$file_format, width = 12, height = 18, bg = "white")


```



