---
title: "Find_markers"
author: "Carla Cohen"
date: "`r Sys.Date()`"
output: html_document
---

Script to find markers on clustered individual samples. 
Follows Cluster.Rmd.
Based on work by Claudia Paul and Adam Cribbs

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)


library(tidyverse)
library(SingleCellExperiment)
library(Seurat)
library(biomaRt)
library(ggplot2)
library(cowplot)
library(clustree)
library(yaml)

#need to update yml file as appropriate
ini <- read_yaml("find_markers_pipeline.yml")


```


Set up which files will be analysed and generate an analysis run directory

```{r}

# list of files to be analysed
# the date of the run needs to be included as a parameter
# the type of normalisation needs to be included
# for other parameters used for clustering, see the readme file in the clustering analysis folder


file_list <- list.files(paste0(ini$cluster_date, "_Clustering.dir/RDS_objects.dir/clustered/", ini$normalisation_method, "/"), pattern="_filtered_clustered_SeuratObject.rds")
names(file_list) <- str_replace(file_list, paste0("_filtered_clustered_SeuratObject.rds"), "")

# make a new output folder for each run, with the date in the directory name
# generate a readme file that shows the parameters used and files analysed

date <- Sys.Date() %>% str_replace_all("-", "")
directory <- paste0(date,"_","Find_markers.dir")
dir.create(directory, showWarnings = FALSE)

sink(file = paste0(directory, "/README.txt"))
cat("This is a description of the following run on ", date, "\n")
cat("\n")
cat("Run Location:\n")
cat(getwd(), "\n")
cat("\n")
cat("This run analysed the following samples:\n")
cat(names(file_list), sep = "\n")
cat("\n")
cat("This run follows the clustering analysis on :\n")
cat(ini$cluster_date, sep = "\n")
cat("\n")
cat("Normalisation method used:\n")
cat(ini$normalisation_method, "\n")
cat("\n")
cat("Clustering resolution used:\n")
cat(ini$resolution, "\n")
cat("\n")
cat("Parameters for FindAllMarkers::\n")
cat("only_pos =", ini$only_pos, "\n")
cat("min_pct =", ini$min_pct, "\n")
cat("logfc_threshold =", ini$logfc_threshold, "\n")
cat("\n")



sink()


```

Read in seurat objects. Each object is from an individual sample and has have normalisation and clustering performed. 

```{r}



#Make a new list for the so objects
so <- list()
# generate a list of sce objects
for (i in 1:length(file_list)){
    print(i)
    so[[i]] <- readRDS(paste0(ini$cluster_date, "_Clustering.dir/RDS_objects.dir/clustered/", ini$normalisation_method, "/", file_list[i]))
    print(file_list[i])
    print(so[[i]])
}


```

### Parameters
Extract some names that depend on the normalisation method performed

```{r}

if (ini$normalisation_method == "SCTransform"){
        
        # save the reduction names
        pca_name <- "sct.pca"
        umap_name <- "sct.umap"
        
        # save the resolution prefix
        res_prefix <- "SCT_snn_res."
        
        # save the assay name
        assay_name <- "SCT"
        
        # save the assay for FindAllMarkers (see below)
        markers_assay <- "RNA"
        
    } else if (ini$normalisation_method == "SCTransformv2"){
    
    # save the reduction names
     pca_name <- "sctv2.pca"
     umap_name <- "sctv2.umap"
     
     # save the resolution prefix
     res_prefix <- "SCTv2_snn_res."
     
     # save the assay name
     assay_name <- "SCTv2"

     # save the assay for FindAllMarkers (see below)
     markers_assay <- "SCTv2"
    
    } else if (ini$normalisation_method == "LogNormalise"){
    
     # save the reduction names
     pca_name <- "lognorm.pca"
     umap_name <- "lognorm.umap"
     
     # save the resolution prefix
     res_prefix <- "RNA_snn_res."
    
     # save the assay name
     assay_name <- "RNA"    
     
     # save the assay for FindAllMarkers (see below)
     markers_assay <- "RNA" 
}
    
print(ini$normalisation_method) 
print(pca_name)
print(umap_name)
print(res_prefix)
print(assay_name)
print(markers_assay)

```

### Normalisation

Log normalisation was already performed in the Filtering workflow. 
Here we need to find variable features and perform scaling for downstream visualisation, if we are working with the SCT assay. 

```{r, message=FALSE}

if (ini$normalisation_method == "SCTransform"){
    for (i in 1:length(so)){
        DefaultAssay(so[[i]]) <- "RNA"
        so[[i]] <- so[[i]] %>% 
           FindVariableFeatures() %>% 
           ScaleData() 
        DefaultAssay(so[[i]]) <- "SCT"
        
    }
}


```



### FindMarkers

Run FindAllMarkers on the RNA assay, if waSCTransform v1 or log normalisation was used.
However, for SCTransformv2 you can run FindAllMarkers on the SCTv2 assay.
Save the output as a text file

```{r, echo=FALSE, message = FALSE}
dir.create(paste0(directory, "/Marker_lists/"), showWarnings = FALSE)
dir.create(paste0(directory, "/Marker_lists/", ini$normalisation_method, "/"), showWarnings = FALSE)

markers <- list()

for (i in 1:length(so)) {
        
        # set chosen resolution (default = 0.5)
        so[[i]][["seurat_clusters"]] <- so[[i]][[paste0(res_prefix, ini$resolution)]]

        # find the markers
        cat("analysing sample", so[[i]]@project.name, "\n")
        markers[[i]] <- FindAllMarkers(so[[i]], 
                                   assay = markers_assay,
                                   only.pos = ini$only_pos, 
                                   min.pct = ini$min_pct, 
                                   logfc.threshold = ini$logfc_threshold)
        
        # save the output
        write.table(markers[[i]], 
                    paste0(directory, "/Marker_lists/", ini$normalisation_method, "/", so[[i]]@project.name, ".txt"), 
                    quote = FALSE, sep = "\t")

}
    

#NB Can read in the text files here if needed

```



Calculate the top 10 markers for each cluster by log2FC

```{r, top10 markers per cluster, echo=FALSE, results='asis'}
top10markers <- list()
for (i in 1:length(so)) {
top10markers[[i]] <- markers[[i]] %>% 
    group_by(cluster) %>% 
    slice_max(n=10, order_by = abs(avg_log2FC))
}

```
Plot the top 10 markers as a dotplot

```{r, general markers, fig.height=60, fig.width=20, echo=FALSE, results='asis'}
plot_list <- list ()
for (i in 1:length(so)) {
    plot_list[[i]] <- DotPlot(so[[i]], 
                              features = unique(top10markers[[i]]$gene),
                              assay = "RNA") + 
         scale_colour_viridis() +
        ggtitle(so[[i]]@project.name) + coord_flip()+
        theme(legend.position="none")
}
title <- ggdraw() + draw_label("Top 10 markers", fontface='bold', size = 24)
p <- plot_grid(plotlist = plot_list, label_size = 12, ncol = 3)
p <- plot_grid(title, p, ncol=1, rel_heights=c(0.05, 1))

p



```

```{r}
dir.create(paste0(directory, "/Marker_dotplots/"))
dir.create(paste0(directory, "/Marker_dotplots/", ini$normalisation_method))
ggsave(paste0(directory, "/Marker_dotplots/", ini$normalisation_method, "/dotplot_top10_markers.", ini$file_type), 
       p, device = ini$file_type, width = 20, height = 45, bg = "white")

```

Use the top 10 genes to plot a heatmap

```{r, general markers, fig.height=60, fig.width=15, echo=FALSE, results='asis'}
plot_list <- list ()
for (i in 1:length(so)) {
    plot_list[[i]] <- DoHeatmap(so[[i]], 
                              features = unique(top10markers[[i]]$gene),
                              assay = "RNA") + 
        #scale_colour_gradientn(colours = viridis(256, option = "D"))+ 
        ggtitle(so[[i]]@project.name)
}
title <- ggdraw() + draw_label("Top 10 markers", fontface='bold', size = 24)
p <- plot_grid(plotlist = plot_list, label_size = 12, ncol = 3)
p <- plot_grid(title, p, ncol=1, rel_heights=c(0.02, 1))

p


```

```{r}
ggsave(paste0(directory, "/Marker_dotplots/", ini$normalisation_method, "/heatmap_top10_markers.", ini$file_type), 
       p, device = ini$file_type, width = 25, height = 45, bg = "white")
```


### Plot groups of known markers

Plots some general markers

```{r, general markers, fig.height=20, fig.width=15, echo=FALSE, results='asis'}
plot_list <- list ()
for (i in 1:length(so)) {
    plot_list[[i]] <- DotPlot(so[[i]], 
                              features =c("COL1A1", "COL1A2", "COL3A1", "DCN", "TPM3", "DES", "PECAM1", "NOTCH3", "PAX7", "PTPRC", "CD163", "CD2", "ADIPOQ", "CD36", "CD69", "TNMD", "TSP4", "TNC","ACAN", "BGN", "SCX"),
                              assay = "RNA") + 
         scale_colour_viridis() +
        ggtitle(so[[i]]@project.name) + coord_flip()
}

title <- ggdraw() + draw_label("General markers", fontface='bold', size = 24)
p <- plot_grid(plotlist = plot_list, label_size = 12, ncol = 2)
p <- plot_grid(title, p, ncol=1, rel_heights=c(0.05, 1))

p



```

```{r}

ggsave(paste0(directory, "/Marker_dotplots/", ini$normalisation_method, "/dotplot_general_markers.", ini$file_type), 
       p, device = ini$file_type, width = 20, height = 28, bg = "white")

```

Plot fibroblast markers

```{r, fibroblast markers, fig.height=20, fig.width=15, echo=FALSE, results='asis'}

plot_list <- list ()
for (i in 1:length(so)) {
    plot_list[[i]] <- DotPlot(so[[i]], 
                              features =c("COL5A1", "LOXL1", "LUM", "FBLN1", "FBLN2", "CD34", "PDGFRA", "ACAN", "BGN", "VCAN", "HSPG2", "SDC4P"),
                              assay = "RNA") + 
         scale_colour_viridis() +
        ggtitle(so[[i]]@project.name) + coord_flip()
}

title <- ggdraw() + draw_label("Fibroblast markers", fontface='bold', size = 24)
p <- plot_grid(plotlist = plot_list, label_size = 12, ncol = 2)
p <- plot_grid(title, p, ncol=1, rel_heights=c(0.05, 1))

p


```

```{r}

ggsave(paste0(directory, "/Marker_dotplots/", ini$normalisation_method, "/dotplot_fibroblast_markers.", ini$file_type), 
       p, device = ini$file_type, width = 20, height = 28, bg = "white")

```

Plot endothelial markers

```{r, endothelial markers, fig.height=20, fig.width=15, echo=FALSE, results='asis'}

plot_list <- list ()
for (i in 1:length(so)) {
    plot_list[[i]] <- DotPlot(so[[i]], 
                              features =c("PECAM1", "SYNPO2", "VCAM1", "FLT1", "KDR", "CD34", "CDH5", "NOTCH3", "ACTA2"),
                              assay = "RNA")  + 
         scale_colour_viridis() +
        ggtitle(so[[i]]@project.name) + coord_flip()
}

title <- ggdraw() + draw_label("Endothelial markers", fontface='bold', size = 24)
p <- plot_grid(plotlist = plot_list, label_size = 12, ncol = 2)
p <- plot_grid(title, p, ncol=1, rel_heights=c(0.05, 1))

p


```

```{r}

ggsave(paste0(directory, "/Marker_dotplots/", ini$normalisation_method, "/dotplot_endothelial_markers.", ini$file_type), 
       p, device = ini$file_type, width = 20, height = 28, bg = "white")

```


Plot muscle markers
```{r, muscle markers, fig.height=20, fig.width=15, echo=FALSE, results='asis'}

plot_list <- list ()
for (i in 1:length(so)) {
    plot_list[[i]] <- DotPlot(so[[i]], 
                              features =c("TPM3", "DES", "TTN", "NOTCH3", "PAX7", "PTPRC", "CD163", "CD2", "ADIPOQ"),
                              assay = "RNA")  + 
         scale_colour_viridis() +
        ggtitle(so[[i]]@project.name) + coord_flip()
}

title <- ggdraw() + draw_label("Muscle markers", fontface='bold', size = 24)
p <- plot_grid(plotlist = plot_list, label_size = 12, ncol = 2)
p <- plot_grid(title, p, ncol=1, rel_heights=c(0.05, 1))

p

```

```{r}

ggsave(paste0(directory, "/Marker_dotplots/", ini$normalisation_method, "/dotplot_muscle_markers.", ini$file_type), 
       p, device = ini$file_type, width = 20, height = 28, bg = "white")

```

Plot mural cell markers

```{r, mural cell markers, fig.height=20, fig.width=15, echo=FALSE, results='asis'}

plot_list <- list ()
for (i in 1:length(so)) {
    plot_list[[i]] <- DotPlot(so[[i]], 
                              features =c("DES", "MCAM", "TAGLN", "NOTCH3", "PDGFRB", "ANPEP"),
                              assay = "RNA")  + 
         scale_colour_viridis() +
        ggtitle(so[[i]]@project.name) + coord_flip()
}

title <- ggdraw() + draw_label("Mural cell markers", fontface='bold', size = 24)
p <- plot_grid(plotlist = plot_list, label_size = 12, ncol = 2)
p <- plot_grid(title, p, ncol=1, rel_heights=c(0.05, 1))

p
```


```{r}
ggsave(paste0(directory, "/Marker_dotplots/", ini$normalisation_method, "/dotplot_mural_markers.", ini$file_type), 
       p, device = ini$file_type, width = 20, height = 28, bg = "white")
```

Plot T cell markers
```{r, T cell markers, fig.height=20, fig.width=15, echo=FALSE, results='asis'}
plots <- list ()
for (i in 1:length(so)) {
    plots[[i]] <- DotPlot(so[[i]], 
                          features =c("CD28", "CD69", "KIT", "CD2", "CD3D", "CD3E", "CD3G", "CD8A", "CD8B"), 
                          assay = "RNA") + 
         scale_colour_viridis() +
        ggtitle(so[[i]]@project.name) + coord_flip()
}

title <- ggdraw() + draw_label("T cell markers", fontface='bold', size = 24)
p <- plot_grid(plotlist = plot_list, label_size = 12, ncol = 2)
p <- plot_grid(title, p, ncol=1, rel_heights=c(0.05, 1))

p
```

```{r}
ggsave(paste0(directory, "/Marker_dotplots/", ini$normalisation_method, "/dotplot_Tcell_markers.", ini$file_type), 
       p, device = ini$file_type, width = 20, height = 28, bg = "white")
```


Plot B cell markers
```{r, B cell markers, fig.height=20, fig.width=15, echo=FALSE, results='asis'}

plots <- list ()
for (i in 1:length(so)) {
    plots[[i]] <- DotPlot(so[[i]], 
                          features =c("MS4A1", "SDC1", "CD37", "BLNK", "CD38", "CD79A", "CD79B", "CD27", "CD78"), 
                          assay = "RNA") + 
         scale_colour_viridis() +
        ggtitle(so[[i]]@project.name) + coord_flip()
}

title <- ggdraw() + draw_label("B cell markers", fontface='bold', size = 24)
p <- plot_grid(plotlist = plot_list, label_size = 12, ncol = 2)
p <- plot_grid(title, p, ncol=1, rel_heights=c(0.05, 1))

p

```
```{r}
ggsave(paste0(directory, "/Marker_dotplots/", ini$normalisation_method, "/dotplot_Bcell_markers.", ini$file_type), 
       p, device = ini$file_type, width = 20, height = 28, bg = "white")
```

Plot Macrophage markers
```{r, Macrophage markers, fig.height=20, fig.width=15, echo=FALSE, results='asis'}

plots <- list ()
for (i in 1:length(so)) {
    plots[[i]] <- DotPlot(so[[i]], 
                          features =c("CD14", "CD16", "CD163", "TREM2", "CX3CR1", "ITGAM", "VSIG4", "MSR1"), 
                          assay = "RNA") + 
         scale_colour_viridis() +
        ggtitle(so[[i]]@project.name) + coord_flip()
}

title <- ggdraw() + draw_label("Macrophage markers", fontface='bold', size = 24)
p <- plot_grid(plotlist = plot_list, label_size = 12, ncol = 2)
p <- plot_grid(title, p, ncol=1, rel_heights=c(0.05, 1))

p

```
```{r}
ggsave(paste0(directory, "/Marker_dotplots/", ini$normalisation_method, "/dotplot_Macrophage_markers.", ini$file_type), 
       p, device = ini$file_type, width = 20, height = 28, bg = "white")
```


#Adipocytes
```{r, Adipocyte markers, fig.height=20, fig.width=15, echo=FALSE, results='asis'}

plots <- list ()
for (i in 1:length(so)) {
    plots[[i]] <- DotPlot(so[[i]], 
                          features =c("ADIPOQ", "FABP4", "LIPE"), 
                          assay = "RNA") + 
         scale_colour_viridis() +
        ggtitle(so[[i]]@project.name) + coord_flip()
}
title <- ggdraw() + draw_label("Adipocyte markers", fontface='bold', size = 24)
p <- plot_grid(plotlist = plot_list, label_size = 12, ncol = 2)
p <- plot_grid(title, p, ncol=1, rel_heights=c(0.05, 1))

p

```

```{r}
ggsave(paste0(directory, "/Marker_dotplots/", ini$normalisation_method, "/dotplot_adipocyte_markers.", ini$file_type), 
       p, device = ini$file_type, width = 20, height = 28, bg = "white")
```

```{r, echo=FALSE}
dir.create (paste0(directory, "/RDS_objects.dir"))
dir.create (paste0(directory, "/RDS_objects.dir/", ini$normalisation_method))


for(i in 1:length(so)){
     saveRDS(so[[i]], 
            paste0(directory, "/RDS_objects.dir/", ini$normalisation_method, "/" , 
                   so[[i]]@project.name, "_decontX_doublet_discard_filtered_clustered_SeuratObject.rds"))
}
```

