---
title: "Achilles Integration"
author: "C.Paul/ Carla Cohen"
date: "08/03/2023"
output: html_document
---

## Set up

Set up and read in the yml file. 

```{r setup, load packages, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(Seurat)
library(patchwork)
library(tidyverse)
library(Matrix)
library(ggplot2)
library(scales)
library(RColorBrewer)
library(harmony)
library(cowplot)
library(clustree)
library(sctransform)
library(scuttle)
library(yaml)

#need to update yml file as appropriate
ini <- read_yaml("integration_pipeline.yml")

```

#### Read in filtered Seurat objects


```{r, read in objects}

#make a list of sample files

file_list <- list.files("RDS_objects.dir/filtered/", pattern=paste0(ini$object_filter, "_SeuratObject.rds"))
names(file_list) <- str_replace(file_list, paste0("_decontX_doublet_", ini$object_filter, "_SeuratObject.rds"), "")


#Make a new list for the seurat objects
so <- list()

# generate a list of Seurat objects
for (i in 1:length(file_list)){
    
    so[[i]] <- readRDS(paste0("RDS_objects.dir/filtered/", file_list[i]))
    #set the project name (should be done in filtering workflow)
    so[[i]]@project.name <- str_replace(file_list[[i]], paste0("_decontX_doublet_", ini$object_filter, "_SeuratObject.rds"), "")
    
}


```

#### Record the parameters used

```{r}
# make a new output folder for each run, with the date in the directory name
# generate a readme file that shows the parameters used and files analysed

date <- Sys.Date() %>% str_replace_all("-", "")
directory <- paste0(date,"_","Integration.dir")
dir.create(directory, showWarnings = FALSE)

sink(file = paste0(directory, "/README.txt"), append = TRUE)
cat("This is a description of the following run on ", date, "\n")
cat("\n")
cat("Run Location:\n")
cat(getwd(), "\n")
cat("\n")
cat("This run analysed the following samples:\n")
cat(names(file_list), sep = "\n")
cat("\n")
cat("Filtering method used:\n")
cat(ini$object_filter, "\n")
cat("\n")
cat("Normalisation method used:\n")
cat(ini$normalisation_method, "\n")
cat("\n")
cat("Number of PCs used:\n")
cat(ini$n_dims, "\n")
cat("\n")
cat("Clustering resolution used:\n")
cat(ini$resolution)

sink()

```


#### Convert rownames of RNA assay data & counts slots to unique names


```{r, include=FALSE}

for (i in 1:length(so)){
  
  # add gene names to RNA assay
  df <- left_join(data.frame("ensembl_gene_id" = rownames(so[[i]]@assays$RNA@data)), so[[i]]@assays$RNA@meta.features)
  rownames(so[[i]]@assays$RNA@data) <- df$unique_name
  rownames(so[[i]]@assays$RNA@counts) <- df$unique_name

}

```

#### Fix the metadata for MSK1250-Ach-MB1
MSK1250-Ach-MB1" has no microanatomical site listed
```{r}
for (i in 1:length(so)){
    print(so[[i]]@project.name)
    print(head(so[[i]][["microanatomical_site"]]))
}


#"MSK1250-Ach-MB1" has no microanatomical site listed
so[[6]][["microanatomical_site"]] <- "MB"

```


#### Parameters
Extract some names that depend on the normalisation method performed

```{r}

if (ini$normalisation_method == "SCTransform"){
        
        # save the reduction names
        pca_name <- "sct.pca"
        umap_name <- "sct.umap"
        
        # save the resolution prefix
        res_prefix <- "SCT_snn_res."
        
    } else if (ini$normalisation_method == "SCTransformv2"){
    
    # save the reduction names
     pca_name <- "sctv2.pca"
     umap_name <- "sctv2.umap"
     
     # save the resolution prefix
     res_prefix <- "SCTv2_snn_res."
    
    } else if (ini$normalisation_method == "LogNormalise"){
    
     # save the reduction names
     pca_name <- "lognorm.pca"
     umap_name <- "lognorm.umap"
     
     # save the resolution prefix
     res_prefix <- "RNA_snn_res."
        
}
    
print(ini$normalisation_method) 
print(pca_name)
print(umap_name)
print(res_prefix)

```


## Merging
Merge the Seurat objects

```{r merge all datasets}

#perform the merge of so1 with all other sos in the list and add cell IDs
so.merge <- merge(so[[1]], y = so[-c(1)], add.cell.ids = names(file_list))

```

#### Perform Normalisation on the merged object

Default method is SCTransform, can also update to logNormalise or SCTransform v2 using the yml file. 
Perform the downstream steps with default parameters and using 30 PCs (this will be optimised later)



```{r, message=FALSE, warning = FALSE}

if (ini$normalisation_method == "SCTransform"){
        
        so.merge <- SCTransform(so.merge,
                           method = "glmGamPoi", #added to increase efficiency
                           vars.to.regress = "percent_mt", 
                           verbose = FALSE) %>% 
        RunPCA(reduction.name = "sct.pca") %>% 
        FindNeighbors(dims = 1:ini$n_dims, reduction = "sct.pca") %>%
        RunUMAP(dims = 1:ini$n_dims, reduction = "sct.pca", reduction.name = "sct.umap") 
    
        # gene names in SCT assay will be the same as RNA assay
        
} else if (ini$normalisation_method == "SCTransformv2"){
    
    so.merge <- SCTransform(so.merge,
                           method = "glmGamPoi", #added to increase efficiency
                           vars.to.regress = "percent_mt", 
                           vst.flavor = "v2", 
                           new.assay.name = "SCTv2", # using SCTransform v2
                           verbose = FALSE) %>% 
        RunPCA(reduction.name = "sctv2.pca") %>% 
        FindNeighbors(dims = 1:ini$n_dims, reduction = "sctv2.pca") %>%
        RunUMAP(dims = 1:ini$n_dims, reduction = "sctv2.pca", reduction.name = "sctv2.umap") 
    
      # # gene names in SCTv2 assay will be the same as RNA assay
        
} else if (ini$normalisation_method == "LogNormalise"){
        so.merge <- so.merge %>% 
           NormalizeData() %>%
           FindVariableFeatures() %>% 
           ScaleData() %>%
           RunPCA(reduction.name = "lognorm.pca")%>%
           FindNeighbors(dims = 1:ini$n_dims, reduction = "lognorm.pca") %>%
           RunUMAP(dims = 1:ini$n_dims, reduction = "lognorm.pca", reduction.name = "lognorm.umap") 
        
    }
    
  
```
#### QC plots on merged object

Elbow Plot

```{r}
p <- ElbowPlot(so.merge, reduction = pca_name, ndims = 50)
p
```



Check distribution by metadata elements on PCA

```{r, by donor, fig.height=10, fig.width=10}
plot_list <- list()

#By sample
plot_list[[1]] <- DimPlot(so.merge, reduction = pca_name, pt.size = .1, group.by = "orig.ident")+
    ggtitle("sample")
#By patient
plot_list[[2]] <- DimPlot(so.merge, reduction = pca_name, pt.size = .1, group.by = "patient")+
    ggtitle ("patient")
#By age
plot_list[[3]] <- DimPlot(so.merge, reduction = pca_name, pt.size = .1, group.by = "age")+
    ggtitle ("age")
#By sex
plot_list[[4]] <- DimPlot(so.merge, reduction = pca_name, pt.size = .1, group.by = "sex")+
    ggtitle ("sex")
#By microanatomy
plot_list[[5]] <- DimPlot(so.merge, reduction = pca_name, pt.size = .1, group.by = "microanatomical_site")+
    ggtitle("microanatomical site")
#By surgical procedure
#plot_list[[6]] <- DimPlot(so.merge, reduction = pca_name, pt.size = .1, group.by = "surgical_procedure")
#By disease status
#plot_list[[7]] <- DimPlot(so.merge, reduction = pca_name, pt.size = .1, group.by = "disease_status")
#By anatomical site
#plot_list[[8]] <- DimPlot(so.merge, reduction = pca_name, pt.size = .1, group.by = "anatomical_site")
#By time to freezing
#plot_list[[9]] <- DimPlot(so.merge, reduction = pca_name, pt.size = .1, group.by = "time_to_freezing")
#By sequencing date
#plot_list[[10]] <- DimPlot(so.merge, reduction = pca_name, pt.size = .1, group.by = "sequencing_date")
#By ethnicity
#plot_list[[11]] <- DimPlot(so.merge, reduction = pca_name, pt.size = .1, group.by = "ethnicity")

title <- ggdraw() + draw_label("Merged data metafeatures on PCA plot", fontface='bold', size = 16)
p <- plot_grid(plotlist = plot_list, ncol = 2) 
p <- plot_grid(title, p, ncol=1, rel_heights=c(0.05, 1))


p
```


```{r, include=FALSE}
#save the plot
dir.create(paste0(directory, "/Integration_Figures.dir/"))
dir.create(paste0(directory, "/Integration_Figures.dir/", ini$normalisation_method))
ggsave(paste0(directory, "/Integration_Figures.dir/", ini$normalisation_method, "/PCA_merged_metafeatures.", ini$file_type), 
       p, device = ini$file_format, width = 10, height = 10, bg = "white")


```


Check distribution by metadata elements on Violin plot

```{r, fig.height=12, fig.width=10}
plot_list <- list()

#By sample
plot_list[[1]] <- VlnPlot(so.merge, pt.size = .1, features = "PC_1", group.by = "orig.ident") +
    ggtitle ("sample")
#By patient
plot_list[[2]] <- VlnPlot(so.merge, pt.size = .1, features = "PC_1", group.by = "patient")+
    ggtitle("patient")
#By age
plot_list[[3]] <- VlnPlot(so.merge, pt.size = .1, features = "PC_1", group.by = "age")+
    ggtitle ("age")
#By sex
plot_list[[4]] <- VlnPlot(so.merge, pt.size = .1, features = "PC_1", group.by = "sex")+
    ggtitle("sex")
#By microanatomy
plot_list[[5]] <- VlnPlot(so.merge, pt.size = .1, features = "PC_1", group.by = "microanatomical_site")+
    ggtitle("microanatomical site")
#By surgical procedure
#plot_list[[6]] <- VlnPlot(so.merge, pt.size = .1, features = "PC_1", group.by = "surgical_procedure")
#By disease status
#plot_list[[7]] <- VlnPlot(so.merge, pt.size = .1, features = "PC_1", group.by = "disease_status")
#By anatomical site
#plot_list[[8]] <- VlnPlot(so.merge, pt.size = .1, features = "PC_1", group.by = "anatomical_site")
#By time to freezing
#plot_list[[9]] <- VlnPlot(so.merge, pt.size = .1, features = "PC_1", group.by = "time_to_freezing")
#By sequencing date
#plot_list[[10]] <- VlnPlot(so.merge, pt.size = .1, features = "PC_1", group.by = "sequencing_date")
#By ethnicity
#plot_list[[11]] <- VlnPlot(so.merge, pt.size = .1, features = "PC_1", group.by = "ethnicity")


title <- ggdraw() + draw_label("Metafeatures on Violin plot", fontface='bold', size = 16)
p <- plot_grid(plotlist = plot_list, ncol = 2) 
p <- plot_grid(title, p, ncol=1, rel_heights=c(0.05, 1))


p


```

```{r, include=FALSE}
#save the plot

ggsave(paste0(directory, "/Integration_Figures.dir/", ini$normalisation_method, "/VlnPlot_merged_metafeatures.", ini$file_type), 
       p, device = ini$file_format, width = 12, height = 10, bg = "white")


```


Visualise contribution of each sample on UMAP

```{r}
p <- DimPlot(so.merge, reduction = umap_name, pt.size = .1, group.by = "orig.ident") +
    ggtitle("Samples on merged UMAP")

p
```

```{r}
ggsave(paste0(directory, "/Integration_Figures.dir/", ini$normalisation_method, "/UMAP_merged_sample.", ini$file_type), 
       p, device = ini$file_format, width = 10, height = 10, bg = "white")

```





## Integration using Harmony

Perform harmony integration

```{r}
so.harmony <- RunHarmony(so.merge, reduction = pca_name, group.by.vars = "orig.ident", plot_convergence = TRUE, dims.use = 1:30)
```

Run UMAP, Find Neighbours and Find Clusters on the integrated object. 
Calculate clustering for a range of resolutsions from 0.1 to 1. 

```{r}
so.harmony <- so.harmony %>% 
    RunUMAP(reduction = "harmony", dims = 1:ini$n_dims, reduction.name = "harmony.umap") %>%
    FindNeighbors(reduction = "harmony", dims = 1:ini$n_dims) %>% 
    FindClusters(resolution = seq(0.1, 1, 0.1))

```

#### QC to check how well the integration has worked

Compare PCA plot from the merged vs integrated data

```{r, fig.width = 14, fig.height=6}

p1 <- DimPlot(so.harmony, reduction = pca_name, group.by = "orig.ident") + ggtitle ("PCA plot merged")
p2 <- DimPlot(so.harmony, reduction = "harmony", group.by = "orig.ident") + ggtitle ("PCA plot integrated")

p <- plot_grid(p1,p2, nrow = 1)
p
```

```{r}
ggsave(paste0(directory, "/Integration_Figures.dir/", ini$normalisation_method, "/PCA_merged_vs_integrated.", ini$file_type), 
       p, device = ini$file_format, width = 14, height = 6, bg = "white")
```


Compare Elbow plot from the merged vs integrated data

```{r}

p1 <- ElbowPlot(so.harmony, ndims = 50, reduction = pca_name) + ggtitle ("Elbow plot merged")
p2 <- ElbowPlot(so.harmony, ndims = 50, reduction = "harmony") + ggtitle ("Elbow plot integrated")
p <- plot_grid(p1,p2)
p
```

```{r}
ggsave(paste0(directory, "/Integration_Figures.dir/", ini$normalisation_method, "/Elbowplot_merged_vs_integrated.", ini$file_type), 
       p, device = ini$file_format, width = 10, height = 6, bg = "white")
```

Which genes are driving each PC?

```{r, fig.width=10, fig.height=10}
p <- DimHeatmap(so.harmony, dims = 1:10, cells = 500, balanced = TRUE, reduction = pca_name)
p
```

```{r}
ggsave(paste0(directory, "/Integration_Figures.dir/", ini$normalisation_method, "/PCA_DimHeatmap.", ini$file_type), 
       p, device = ini$file_format, width = 10, height = 10, bg = "white")
```


Plot the clustree to determine the most appropriate resolution to use

```{r, fig.height= 10, fig.width=10}

p <- clustree(so.harmony, prefix = res_prefix) + 
        ggtitle("Clustree")
    
p

```


```{r, include=FALSE}
#save the plot
ggsave(paste0(directory, "/Integration_Figures.dir/", ini$normalisation_method, "/Clustree_plot.", ini$file_type), 
       p, device = ini$file_format, width = 10, height = 10, bg = "white")


```

Plot UMAPs at multiple clustering resolutions

```{r, fig.width=10, fig.height=16}

resolutionList <- grep("_snn_res", colnames(so.harmony@meta.data), value = TRUE)
plot_list <- list()

for (resolution in resolutionList){
      plot_list [[resolution]] <- DimPlot(object = so.harmony, label = TRUE, reduction = "harmony.umap", group.by = resolution)+
          theme(legend.position="none")
      }

title <- ggdraw() + draw_label("UMAPs of clustering resolutions 0.1-1", fontface='bold', size = 24)
p <- plot_grid(plotlist = plot_list, ncol = 3) 
p <- plot_grid(title, p, ncol=1, rel_heights=c(0.05, 1))


p

```

```{r, include=FALSE}
#save the plot
ggsave(paste0(directory, "/Integration_Figures.dir/", ini$normalisation_method, "/UMAP_compare_resolutions.", ini$file_type), 
       p, device = ini$file_format, width = 10, height = 16, bg = "white")


```

Select a resolution and set seurat clusters to that resolution 

```{r} 

so.harmony[["seurat_clusters"]] <- so.harmony[[paste0(res_prefix, ini$resolution)]]
Idents(so.harmony) <- so.harmony[[paste0(res_prefix, ini$resolution)]]

```

#### Cell cycle

Look at cell cycle markers using CellCycleScoring

First calculate the phase for each cell, using an inbuilt database (in Seurat) of known cell cycle marker genes. 

```{r}

#Run CellCycleScoring algorithm, results will appear in the metadata 
so.harmony <- CellCycleScoring(so.harmony, s.features = cc.genes$s.genes, g2m.features = cc.genes$g2m.genes)

# view cell cycle scores and phase assignments
head(so.harmony[[]])

```

Visualise the output of cell cycle markers

```{r, fig.width=16, fig.height=10}
# Visualize the distribution of cell cycle markers 
p1 <- RidgePlot(so.harmony, features = c("PCNA", "TOP2A", "MCM6", "MKI67"), group.by = "Phase", ncol = 4)
title <- ggdraw() + draw_label("Expression of cell cycle marker genes", fontface='bold', size = 16)


# Visualise cell cycle assignment on PCA
p2 <- DimPlot(so.harmony, reduction = "sct.pca", group.by = "Phase")+
    ggtitle("Phase mapped on PCA")

# Visualise cell cycle assignment on UMAP
p3 <- DimPlot(so.harmony, reduction = "harmony.umap", group.by = "Phase", pt.size = 0.1)+
    ggtitle("Phase mapped on UMAP")

p4 <- plot_grid(p2, p3)

p <- plot_grid(title, p1, p4, nrow = 3, rel_heights=c(0.1, 0.6, 1))

p
```
Cell cycle could be regressed out in later steps if required. 

```{r, include=FALSE}
#save the plot
ggsave(paste0(directory, "/Integration_Figures.dir/", ini$normalisation_method, "/Cell_cycle_metrics.", ini$file_type), 
       p, device = ini$file_type, width = 16, height = 10, bg = "white")


```

## Save objects 
Save merged & integrated Seurat objects
```{r}
dir.create(paste0(directory, "/RDS_objects.dir/"))
dir.create(paste0(directory, "/RDS_objects.dir/", ini$normalisation_method, "/"))

saveRDS(so.merge, paste0(directory, "/RDS_objects.dir/", ini$normalisation_method, "/Achilles_merge_SeuratObject.rds"))
saveRDS(so.harmony, paste0(directory, "/RDS_objects.dir/", ini$normalisation_method, "/Achilles_harmony_SeuratObject.rds"))

```


## Integration visualisation
Visualization of merged and integrated data via UMAP split by sample

```{r, fig.width = 12}
p1 <- DimPlot(so.merge, reduction = umap_name, pt.size = .1, group.by = "orig.ident") +
    ggtitle("Merged")
p2 <- DimPlot(so.harmony, reduction = "harmony.umap", pt.size = .1, group.by = "orig.ident")+ 
    ggtitle ("Harmony Integration")
p <- plot_grid (p1, p2)
p

```

```{r, include=FALSE}
#save the plot
ggsave(paste0(directory, "/Integration_Figures.dir/", ini$normalisation_method, "/UMAP_merged_vs_harmony.", ini$file_type), 
       p, device = ini$file_format, width = 14, height = 6, bg = "white")


```

Visualization of integrated data via UMAP at chosen resolution

```{r, fig.height=6, fig.width=6}
p <- DimPlot(so.harmony, reduction = "harmony.umap", pt.size = .1)+ 
    ggtitle (paste0("Harmony Integration Clustering at resolution = ", ini$resolution))
p


```

```{r, include=FALSE}
#save the plot
ggsave(paste0(directory, "/Integration_Figures.dir/", ini$normalisation_method, 
              "/UMAP_harmony_clustering_res_", ini$resolution, ".", ini$file_type), 
       p, device = ini$file_format, width = 10, height = 10, bg = "white")


```



UMAP of integrated data by sample with each sample on a separate plot

```{r, fig.width=14}

p <- DimPlot(so.harmony, reduction = "harmony.umap", split.by = "orig.ident", ncol = 5)

p
```

```{r, include=FALSE}
#save the plot
ggsave(paste0(directory, "/Integration_Figures.dir/", ini$normalisation_method, "/UMAP_harmony_by_sample.", ini$file_type), 
       p, device = ini$file_format, width = 14, height = 6, bg = "white")


```

Check distribution by metadata elements on UMAP following integration

```{r, by donor, fig.height=10, fig.width=10}
plot_list <- list()

#By sample
plot_list[[1]] <- DimPlot(so.harmony, reduction = "harmony.umap", pt.size = .1, group.by = "orig.ident")+
    ggtitle("sample")
#By patient
plot_list[[2]] <- DimPlot(so.harmony, reduction = "harmony.umap", pt.size = .1, group.by = "patient")+
    ggtitle ("patient")
#By age
plot_list[[3]] <- DimPlot(so.harmony, reduction = "harmony.umap", pt.size = .1, group.by = "age")+
    ggtitle ("age")
#By sex
plot_list[[4]] <- DimPlot(so.harmony, reduction = "harmony.umap", pt.size = .1, group.by = "sex")+
    ggtitle ("sex")
#By microanatomy
plot_list[[5]] <- DimPlot(so.harmony, reduction = "harmony.umap", pt.size = .1, group.by = "microanatomical_site")+
    ggtitle("microanatomical site")
#By surgical procedure
#plot_list[[6]] <- DimPlot(so.harmony, reduction = "harmony.umap", pt.size = .1, group.by = "surgical_procedure")
#By disease status
#plot_list[[7]] <- DimPlot(so.harmony, reduction = "harmony.umap", pt.size = .1, group.by = "disease_status")
#By anatomical site
#plot_list[[8]] <- DimPlot(so.harmony, reduction = "harmony.umap", pt.size = .1, group.by = "anatomical_site")
#By time to freezing
#plot_list[[9]] <- DimPlot(so.harmony, reduction = "harmony.umap", pt.size = .1, group.by = "time_to_freezing")
#By sequencing date
#plot_list[[10]] <- DimPlot(so.harmony, reduction = "harmony.umap", pt.size = .1, group.by = "sequencing_date")
#By ethnicity
#plot_list[[11]] <- DimPlot(so.harmony, reduction = "harmony.umap", pt.size = .1, group.by = "ethnicity")

title <- ggdraw() + draw_label("Metafeatures on UMAP after harmony integration", fontface='bold', size = 24)
p <- plot_grid(plotlist = plot_list, ncol = 2) 
p <- plot_grid(title, p, ncol=1, rel_heights=c(0.05, 1))


p
```

```{r, include=FALSE}
#save the plot
ggsave(paste0(directory, "/Integration_Figures.dir/", ini$normalisation_method, "/UMAP_harmony_metafeatures.", ini$file_type), 
       p, device = ini$file_format, width = 10, height = 10, bg = "white")


```

Plot metafeatures on Vln plot of harmony principal component

```{r, fig.height=10, fig.width=12}
plot_list <- list()

#By sample
plot_list[[1]] <- VlnPlot(object = so.harmony, features = "harmony_1", group.by = "orig.ident", pt.size = .1) +
    ggtitle ("sample")
#By patient
plot_list[[2]] <- VlnPlot(so.harmony, pt.size = .1, features = "harmony_1", group.by = "patient")+
    ggtitle("patient")
#By age
plot_list[[3]] <- VlnPlot(so.harmony, pt.size = .1, features = "harmony_1", group.by = "age")+
    ggtitle ("age")
#By sex
plot_list[[4]] <- VlnPlot(so.harmony, pt.size = .1, features = "harmony_1", group.by = "sex")+
    ggtitle("sex")
#By microanatomy
plot_list[[5]] <- VlnPlot(so.harmony, pt.size = .1, features = "harmony_1", group.by = "microanatomical_site")+
    ggtitle("microanatomical site")
#By surgical procedure
#plot_list[[6]] <- VlnPlot(so.harmony, pt.size = .1, features = "harmony_1", group.by = "surgical_procedure")
#By disease status
#plot_list[[7]] <- VlnPlot(so.harmony, pt.size = .1, features = "harmony_1", group.by = "disease_status")
#By anatomical site
#plot_list[[8]] <- VlnPlot(so.harmony, pt.size = .1, features = "harmony_1", group.by = "anatomical_site")
#By time to freezing
#plot_list[[9]] <- VlnPlot(so.harmony, pt.size = .1, features = "harmony_1", group.by = "time_to_freezing")
#By sequencing date
#plot_list[[10]] <- VlnPlot(so.harmony, pt.size = .1, features = "harmony_1", group.by = "sequencing_date")
#By ethnicity
#plot_list[[11]] <- VlnPlot(so.harmony, pt.size = .1, features = "harmony_1", group.by = "ethnicity")


title <- ggdraw() + draw_label("Metafeatures on Violin plot after Harmony integration", fontface='bold', size = 16)
p <- plot_grid(plotlist = plot_list, ncol = 2) 
p <- plot_grid(title, p, ncol=1, rel_heights=c(0.05, 1))


p


```

```{r, include=FALSE}
#save the plot
ggsave(paste0(directory, "/Integration_Figures.dir/", ini$normalisation_method, "/VlnPlot_harmony_metafeatures.", ini$file_type), 
       p, device = ini$file_format, width = 12, height = 10, bg = "white")


```


Plot UMAP of microanatomical site on separate plots


```{r, fig.width = 10, fig.height = 10}

p <- DimPlot(so.harmony, reduction = "harmony.umap", pt.size = .1, split.by = "microanatomical_site", ncol = 2)+
    ggtitle("microanatomical site")
p
```

```{r, include=FALSE}
#save the plot
ggsave(paste0(directory, "/Integration_Figures.dir/", ini$normalisation_method, "/UMAP_harmony_by_microanatomy.", ini$file_type), 
       p, device = ini$file_type, width = 10, height = 10, bg = "white")


```

Plot QC metrics on VlnPlot

```{r, fig.height=6, fig.width=16}
p1<- VlnPlot(so.harmony, features = "nFeature_RNA")
p2<- VlnPlot(so.harmony, features = "nCount_RNA")
p3<- VlnPlot(so.harmony, features = "percent_mt")
p <- plot_grid(p1, p2, p3, nrow = 1)
p

```

```{r, include=FALSE}
#save the plot
ggsave(paste0(directory, "/Integration_Figures.dir/", ini$normalisation_method, "/VlnPlot_harmony_QC_metrics.", ini$file_type), 
       p, device = ini$file_type, width = 10, height = 10, bg = "white")


```

Plot QC metrics on UMAP

```{r, fig.height=6, fig.width=16}
p1<- FeaturePlot(so.harmony, features = "nFeature_RNA")
p2<- FeaturePlot(so.harmony, features = "nCount_RNA")
p3<- FeaturePlot(so.harmony, features = "percent_mt")
p <- plot_grid(p1, p2, p3, nrow = 1)
p

```

```{r, include=FALSE}
#save the plot
ggsave(paste0(directory, "/Integration_Figures.dir/", ini$normalisation_method, "/UMAP_harmony_QC_metrics.", ini$file_type), 
       p, device = ini$file_type, width = 10, height = 10, bg = "white")


```



