---
title: "Microanatomy comparison"
author: "Carla Cohen"
date: "`r Sys.Date()`"
output: html_document
---

The aim of this notebook is to compare the achilles tendon microanatomies. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(Seurat)
library(cowplot)
library(yaml)
library(viridis)

ini <- read_yaml("microanatomy_pipeline.yml")

```

Set up

```{r}

# list of files to be analysed
# the date of the run needs to be included as a parameter
# the type of normalisation needs to be included

file_list <- list.files(paste0(ini$annotation_date, "_Annotation.dir/RDS_objects.dir/", ini$normalisation_method, "/"), pattern=".rds")
names(file_list) <- str_replace(file_list, ".rds", "")

# make a new output folder for each run, with the date in the directory name
# generate a readme file that shows the parameters used and files analysed

date <- Sys.Date() %>% str_replace_all("-", "")
directory <- paste0(date,"_","Microantomy.dir")
dir.create(directory, showWarnings = FALSE)

sink(file = paste0(directory, "/README.txt"))
cat("This is a description of the following run on ", date, "\n")
cat("\n")
cat("Run Location:\n")
cat(getwd(), "\n")
cat("\n")
cat("This run analysed the following samples:\n")
cat(names(file_list), sep = "\n")
cat("\n")
cat("This run follows the annotation analysis on :\n")
cat(ini$annotation_date, sep = "\n")
cat("\n")
cat("Normalisation method used:\n")
cat(ini$normalisation_method, "\n")
cat("\n")
cat("Clustering resolution used:\n")
cat(ini$resolution, "\n")
cat("\n")

sink()


```

Read in the integrated Seurat object

```{r}

so.harmony <- readRDS(paste0(ini$annotation_date, "_Annotation.dir/RDS_objects.dir/", ini$normalisation_method, "/", file_list[1]))
so.harmony

```

Plot the UMAP separated by microanatomy

```{r, fig.height=10, fig.width=14}

p <- DimPlot(so.harmony, reduction = "harmony.umap", label = TRUE, split.by = "microanatomical_site", ncol = 2)
p
```
Question: what are the number of cells in each cluster for each microanatomy?

```{r}
#create a df with the relevant info
df <- so.harmony[[]] %>% dplyr::select (microanatomical_site, seurat_clusters)
df$identity <- Idents(so.harmony)
df

# create a second df with numbers of cells per cluster per microanatomy
cell_numbers <- data.frame()

for (level in levels(df$seurat_clusters)){
    
    enth <- df %>% filter(seurat_clusters == level) %>% filter(microanatomical_site == "Enth") %>% nrow()
    mb <- df %>% filter(seurat_clusters == level) %>% filter(microanatomical_site == "MB") %>% nrow()
    mtj <- df %>% filter(seurat_clusters == level) %>% filter(microanatomical_site == "MTJ") %>% nrow()
    muscle <- df %>% filter(seurat_clusters == level) %>% filter(microanatomical_site == "muscle") %>% nrow()
    total <- df %>% filter(seurat_clusters == level) %>% nrow()
    ident <- df %>% filter(seurat_clusters == level) %>% slice_head(n=1) %>% pull (identity)

    column_number <- as.numeric(level) + 1

    cell_numbers[column_number, 1] <- level
    cell_numbers[column_number, 2] <- enth
    cell_numbers[column_number, 3] <- mb
    cell_numbers[column_number, 4] <- mtj
    cell_numbers[column_number, 5] <- muscle
    cell_numbers[column_number, 6] <- total
    cell_numbers[column_number, 7] <- ident

    
}

colnames(cell_numbers) <- c("cluster", "Enth", "MB", "MTJ", "Muscle", "Total", "Identity")

cell_numbers

```
 
 Plot the total number of cells in each cluster
```{r}
ggplot(cell_numbers, aes (x = Identity, y = Total))+
    geom_col()+
    ggtitle("Number of cells per cluster")+
    theme(plot.title = element_text(size = 24))+
    theme_cowplot()+
    theme(axis.text.x = element_text(angle = 45, hjust=1))+
    theme(axis.title.x = element_blank())
```
 
 
 
 
 Tidy the data!
 
```{r}
cell_number_tidy <- cell_numbers %>% 
    pivot_longer(2:5, values_to = "number_of_cells", names_to = "microanatomy")
cell_number_tidy
```
 


Plot

```{r, fig.width = 10, fig.height=6}
p <- ggplot(cell_number_tidy, aes(x = Identity, y = number_of_cells, fill = microanatomy))+
    geom_col(position = "dodge")+
    ggtitle("Number of cells per cluster by microanatomy")+
    theme(plot.title = element_text(size = 24))+
    scale_fill_viridis_d(begin = 0, end = 0.75)+
    theme_cowplot()+
    theme(axis.text.x = element_text(angle = 45, hjust=1))
```

Make a df with cell fraction per microanatomy

```{r}
cell_fractions <- cell_numbers

cell_fractions$Enth_fraction <- cell_fractions$Enth/sum(cell_fractions$Enth)
cell_fractions$MB_fraction <- cell_fractions$MB/sum(cell_fractions$MB)
cell_fractions$MTJ_fraction <- cell_fractions$MTJ/sum(cell_fractions$MTJ)
cell_fractions$Muscle_fraction <- cell_fractions$Muscle/sum(cell_fractions$Muscle)

cell_fractions <- cell_fractions %>% dplyr::select(!c("Enth", "MB", "MTJ", "Muscle", "Total"))
cell_fractions
```
Tidy the data!
 
```{r}
cell_fraction_tidy <- cell_fractions %>% 
    pivot_longer(3:6, values_to = "fraction_of_cells", names_to = "microanatomy")
cell_fraction_tidy
```
Plot the fraction of cells per cluster by microanatomy

```{r, fig.width=10, fig.height=6}
p <- ggplot(cell_fraction_tidy, aes(x = Identity, y = fraction_of_cells, fill = microanatomy))+
    geom_col(position = "dodge")+
    ggtitle("Fraction of cells per cluster by microanatomy")+
    theme(plot.title = element_text(size = 24))+
    scale_fill_viridis_d(begin = 0, end = 0.75)+
    theme_cowplot()+
    theme(axis.text.x = element_text(angle = 45, hjust=1))
p
```

